---
title: "H1 Exploration"
author: "Anonymous"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(janitor)
```


# Purpose 

This Rmarkdown will be used to explore preliminary outputs of behavior space H1:
"what is the relationship between CPUE and estimated density metrics from other methods?"

# Data

I used two sources of data. 

The first one, `H1...csv` is the output of the behavior space experiment for H1. It includes all four sampling methods, four different mussel densities, and a clustered and random spatial distribution of mussels. There are 100 iterations per each scenario

The second data source, `H1_just_CPUE....csv` is from the H1_just_CPUE behavior space experiment. It drills down on just the random walk sampling method. It includes one sampling method, random-walk, mussel densities between 0 and 10 at increments of 0, and random and clustered spatial distributions. There are 100 iterations per each scenario.

# Analysis of Sampling Methods

```{r}
#read in data (for now copied into R project)
raw_data <- read.csv("data/H1_6-17-2024.csv", sep = ",")

#clean names
clean_data <- clean_names(raw_data)
```

In this section I will examine the random walk sampling method

```{r}
#filter for random walk
random_walk <- clean_data %>% 
  filter(sampling_method == "timed-search")

#convert true mussel density to a factor
random_walk$true_mussel_density <- as.factor(random_walk$true_mussel_density)

#convert cpue to numeric
random_walk$cpue <- as.numeric(random_walk$mussels_per_person_hour)

#plot boxplot
ggplot(data = random_walk, 
       aes(x = true_mussel_density, y = cpue, color = spatial_distribution)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +
  labs(title = "Random Walk CPUE at different densities boxplot")

#plot jitter plot
ggplot(data = random_walk, 
       aes(x = true_mussel_density, y = cpue, color = spatial_distribution)) +
  geom_jitter() +
  labs(title = "Random Walk CPUE at different densities jitter plot")
```

In this section I will examine the simple random sample (SRS) Sampling Method

```{r}
#filter for simple random sample
srs <- clean_data %>% 
  filter(sampling_method == "simple-random-sample")

#convert mussel density to factor
srs$true_mussel_density <- as.factor(srs$true_mussel_density)

#convert estimated density to numeric
srs$estimated_density <- as.numeric(srs$estimated_density)

#boxplot
ggplot(data = srs, 
       aes(x = true_mussel_density, y = estimated_density, 
           color = spatial_distribution)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +
  labs(title = "SRS estimated density at different densities boxplot")
```


```{r}
#calculate variance per density and distribution group
var_est_density <- clean_data %>% 
  filter(sampling_method == "adaptive-cluster") %>% 
  group_by(true_mussel_density, spatial_distribution) %>% 
  summarise(var_est_density = var(hh_metric))

ggplot(data = var_est_density, 
       aes(x = as.factor(true_mussel_density), y = var_est_density, 
           color = spatial_distribution)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +
  labs(title = "ACS variance in estimated density at different densities boxplot")
```

In this section I will examine the transect Sampling Method

```{r}
#filter for transect
transect <- clean_data %>% 
  filter(sampling_method == "transect")

#convert mussel density to factor
transect$true_mussel_density <- as.factor(transect$true_mussel_density)

#convert estimated density to numeric
transect$estimated_density <- as.numeric(transect$estimated_density)

#create a boxplot
ggplot(data = transect, 
       aes(x = true_mussel_density, y = estimated_density, 
           color = spatial_distribution)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +
  labs(title = "Transect estimated density at different densities boxplot")
```

In this section I will examine the Adaptive Cluster Sampling (ACS) Method


```{r}
#filter for adaptive cluster
acs <- clean_data %>% 
  filter(sampling_method == "adaptive-cluster")

#convert mussel density to factor
acs$true_mussel_density <- as.factor(acs$true_mussel_density)

#convert hh metric to numeric
acs$hh_metric <- as.numeric(acs$hh_metric)

#make a boxplot
ggplot(data = acs, 
       aes(x = true_mussel_density, y = hh_metric, 
           color = spatial_distribution)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +
  labs(title = "ACS HH at different densities boxplot")
```

# How does CPUE relate to true density?

calculate co-variance and pearson's correlation coefficient between true density and CPUE.

```{r}
#using the random walk subset created above...

#convert mussel density to numeric
random_walk$true_mussel_density <- as.numeric(random_walk$true_mussel_density)

#covariance and pearson correlation coefficient
random_walk %>% 
  group_by(spatial_distribution) %>% 
  summarise(
    covarience = cov(true_mussel_density, cpue),
    correlation_coef = cor(true_mussel_density, cpue, method = "pearson"))
```

Test for a linear relationship 

```{r}
#create a linear model linear relationship
lm <- lm(true_mussel_density~cpue, data = random_walk)

#summarise to view important metrics
summary(lm)
```

In conclusion, yes true mussel density and CPUE are correlated. There is a decent linear relationship between the two metrics.

# Taking a Deeper Dive into CPUE 

I reran the simulation with just random walking, and densities between 0 and 10 at increments of 0.1

```{r}
#read in new data
just_cpue_raw <- read.csv("data/H1_just_CPUE_6-17-2024.csv", sep = ",")

#clean names
just_cpue <- clean_names(just_cpue_raw)
```

Make some exploratory plots

```{r}
#create a geom point to look at data
ggplot(data = just_cpue, 
       aes(x = true_mussel_density, y = mussels_per_person_hour, color = spatial_distribution)) +
  geom_point() +
  labs(title = "CPUE vs true mussel density")

#let's look at CPUE at lower density..
#filter for low density
low_density <- just_cpue %>% 
  filter(true_mussel_density < 1.5)

ggplot(data = low_density, 
       aes(x = true_mussel_density, y = mussels_per_person_hour, color = spatial_distribution)) +
  geom_point() +
  labs(title = "CPUE vs true mussel density just low density")

#let's ignore the variation and just look at mean cpue per density
#group by mussel density (remember there were 100 iterations per density)
#calculate mean CPUE
means <- just_cpue %>% 
  group_by(true_mussel_density, spatial_distribution) %>% 
  summarise(mean_cpue = mean(mussels_per_person_hour))

ggplot(data = means, 
       aes(x = true_mussel_density, y = mean_cpue,
           color = spatial_distribution)) +
  geom_point() +
  labs(title = "Mean CPUE vs true mussel density")

#now look at mean CPUE at low density...is there a faint curve?
low_density_means <- means %>% 
  filter(true_mussel_density < 1.5)

ggplot(data = low_density_means, 
       aes(x = true_mussel_density, y = mean_cpue, 
           color = spatial_distribution)) +
  geom_point() +
  labs(title = "Mean CPUE vs true mussel density just low density")
```

I also want to look at the variance in CPUE across different true densities

```{r}
#calculate variance per density and distribution group
var_cpue <- just_cpue %>% 
  group_by(true_mussel_density, spatial_distribution) %>% 
  summarise(variance_cpue = var(mussels_per_person_hour))

#plot variance in cpue over different density
ggplot(data = var_cpue, 
       aes(x = true_mussel_density, y = variance_cpue,
           color = spatial_distribution)) +
  geom_line() +
  labs(title = "Variance in CPUE vs true mussel density")
  
```

# Is the relationship between CPUE and density linear??

compare linear regression model to log regression model

```{r}
#create linear regression model
lm <- lm(mussels_per_person_hour ~ true_mussel_density, data = just_cpue)

#summarize to view important metrics
summary(lm)
```


```{r}
plot(just_cpue$true_mussel_density, just_cpue$mussels_per_person_hour)

#define x-values to use for regression line
true_mussel_density = seq(from=0,to=10,length.out=1000)

#use the model to predict the y-values based on the x-values
cpue = predict(lm, 
               newdata=list(true_mussel_density=seq(from=0.01,to=10,length.out=1000)), 
          
               interval="confidence")

cpue_df <- as.data.frame(cpue)

#add the fitted regression line to the plot (lwd specifies the width of the line)
matlines(true_mussel_density, cpue, lwd=2)

ggplot() +
  geom_point(color = "blue", data = just_cpue, 
             aes(x = true_mussel_density, y = mussels_per_person_hour)) +
  geom_abline(slope = coef(lm)[["true_mussel_density"]], 
              intercept = coef(lm)[["(Intercept)"]]) +
  labs(title = "Relationship between mussels per person-hour and true mussel density",
       x = "mussels per square meter",
       y = "mussels per\nperson-hour") +
  theme_bw() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
  

```


```{r}
#first filter out 0 values 
no_zero <- just_cpue %>% 
  filter(mussels_per_person_hour > 0)

#log regression
log <- lm(formula = mussels_per_person_hour ~ log(true_mussel_density), data = no_zero)

#print summary
summary(log)
```


```{r}
plot(just_cpue$true_mussel_density, just_cpue$mussels_per_person_hour)

#define x-values to use for regression line
true_mussel_density = seq(from=0,to=10,length.out=1000)

#use the model to predict the y-values based on the x-values
cpue = predict(log, 
               newdata=list(true_mussel_density=seq(from=0.01,to=10,length.out=1000)), 
          
               interval="confidence")

#add the fitted regression line to the plot (lwd specifies the width of the line)
matlines(true_mussel_density,cpue, lwd=2)

```

the previous models included both clustered and random spatial distributions. Would the linear model improve if we only had one type of spatial distribution?

```{r}
#filter for random spatial distribution
just_cpue_random <- filter(just_cpue, spatial_distribution == "random")

#filter for just clustered distribution
just_cpue_clustered <- filter(just_cpue, spatial_distribution == "Clumped-Poisson")

#create linear regression model for random
lm_random <- lm(mussels_per_person_hour ~ true_mussel_density, data = just_cpue_random)

#create linear regression model for clustered
lm_cluster <- lm(mussels_per_person_hour ~ true_mussel_density, data = just_cpue_clustered)

#summarize to view important metrics
summary(lm_random)
summary(lm_cluster)
```

Unsurprisingly, we can see when mussels are randomly distributed, the cpue has a linear relationship with cpue and even more of the variance can be explained by the model (indicated by the higher R-squared value).

# Quadrats Sampled/Time Searching vs Density

Now I will look at how much area/time is covered per random-walk model run. Area covered can be examined via `total_quadrats_sampled` but it's important to remember that a surveyor can re-search a quadrat that has already been sampled and this will *not* be reflected in the `total_quadrats_sampled` variable. Also, remember there are always three surveyors and that in this model simulation quadrats are (0.5m X 0.5m)

```{r}
#convert total quadrats to numeric
just_cpue$total_quadrats_sampled <- as.numeric(just_cpue$total_quadrats_sampled)

#covert total quadrats to square m
just_cpue_m <- just_cpue %>% 
  mutate(m_sampled = total_quadrats_sampled * (0.5 * 0.5))

#plot m sampled
ggplot(data = just_cpue_m, 
       aes(y = m_sampled, x = true_mussel_density, color = spatial_distribution)) +
  geom_point() +
  labs(title = "square meters sampled by all surveyors vs mussel density")

#plot average m per person sampled
ggplot(data = just_cpue_m, 
       aes(y = m_sampled/3, x = true_mussel_density, color = spatial_distribution)) +
  geom_point() +
  labs(title = "square meters sampled per surveyor vs mussel density")
```

let's also look at variance in number of mussels detected per density

```{r}
#convert mussels detected to numeric
just_cpue$total_mussels_detected <- as.numeric(just_cpue$total_mussels_detected)

#plot mussels detected 
ggplot(data = just_cpue, 
       aes(y = total_mussels_detected, x = true_mussel_density, 
           color = spatial_distribution)) +
  geom_point() +
  labs(title = "Mussels Detected vs mussel density")

#variation
var_m_detect <- just_cpue %>% 
  group_by(true_mussel_density, spatial_distribution) %>% 
  summarise(var_mussels_detected = var(total_mussels_detected))

#plot mussels detected 
ggplot(data = var_m_detect, 
       aes(y = var_mussels_detected, x = true_mussel_density, 
           color = spatial_distribution)) +
  geom_line() +
  labs(title = "Variance in Mussels Detected (timed search)",
       x = "mussels per square meter",
       y = "variance in number of\nmussels detected",
       color = "mussel distribution") +
  theme_bw() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
```

