---
title: "H1 Timed Search"
author: "Anonymous"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(nlme)
library(cowplot)
```

# Purpose 

This Rmarkdown will be used to explore preliminary outputs of behavior space H1:
"what is the relationship between CPUE and true density?"

# Data

Timed search was ran for 4.5 person hours when mussels had seven spatial distributions: random, matern clumps (subdivide up into five distibutions), and poisson clumps with 5 clumps and mean of 3. Density ranged from 0.0 to 7 with increments of 0.1. there were 100 iterations for each scenario

```{r}
#read in data
matern <- read.csv("data/H1_matern_timed.csv") %>% 
  clean_names() %>% 
  filter(detectability_of_common_mussels == 0.3)

poisson <- read.csv("data/H1_poisson_timed.csv") %>% 
  clean_names() %>% 
  filter(detectability_of_common_mussels == 0.3)

random <- read.csv("data/H1_random_timed.csv") %>% 
  clean_names() %>% 
  filter(detectability_of_common_mussels == 0.3)

os_matern <- matern %>% 
  filter(matern_clump_placement == "One shore only")

os_om_matern <- matern %>% 
  filter(matern_clump_placement == "One shore; one middle")

os_om_two_matern <- matern %>% 
  filter(matern_clump_placement == "One shore; one middle; 2 additional")

os_two_matern <- matern %>% 
  filter(matern_clump_placement == "One shore; 2 additional")

three_matern <- matern %>% 
  filter(matern_clump_placement == "Randomly placed")
```

# linear models

```{r}
#create linear regression models
lm_random <- lm(mussels_per_person_hour ~ true_mussel_density, data = random)


lm_poisson <- lm(mussels_per_person_hour ~ true_mussel_density, data = poisson)


lm_os <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_matern)


lm_os_om <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_om_matern)


lm_os_om_two <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_om_two_matern)


lm_os_two <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_two_matern)


lm_three <- lm(mussels_per_person_hour ~ true_mussel_density, data = three_matern)


```

# test for heteroscedasticity 

```{r}
#yes, data is heteroscedastic
plot(fitted(lm_random), resid(lm_random), xlab = "fitted vals", ylab = "residuals")
```

# test for normality

```{r}
#qqplot
qqnorm(residuals(lm_random))
qqline(residuals(lm_random), col = "red")

#histogram
hist(residuals(lm_random), breaks = 50)

#data is normal enough
```

# weighted regressions

```{r}
#create weights 
w_random <- 1/abs(residuals(lm_random))

#weighted regression
wlm_random <- lm(mussels_per_person_hour ~ true_mussel_density, data = random, weights = w_random)

#testing 
#plot(wlm_random)
summary(wlm_random)
```


```{r}
w_poisson <- 1/abs(residuals(lm_poisson))
wlm_poisson <- lm(mussels_per_person_hour ~ true_mussel_density, data = poisson, weights = w_poisson)
summary(wlm_poisson)
```


```{r}
w_os <- 1/abs(residuals(lm_os))
wlm_os <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_matern, weights = w_os)
summary(wlm_os)
```


```{r}
w_os_om <- 1/abs(residuals(lm_os_om))
wlm_os_om <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_om_matern, weights = w_os_om)
summary(wlm_os_om)
```


```{r}
w_os_om_two <- 1/abs(residuals(lm_os_om_two))
wlm_os_om_two <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_om_two_matern, weights = w_os_om_two)
summary(wlm_os_om_two)
```


```{r}
w_os_two <- 1/abs(residuals(lm_os_two))
wlm_os_two <- lm(mussels_per_person_hour ~ true_mussel_density, data = os_two_matern, weights = w_os_two)
summary(wlm_os_two)
```


```{r}
w_three <- 1/abs(residuals(lm_three))
wlm_three <- lm(mussels_per_person_hour ~ true_mussel_density, data = three_matern, weights = w_three)
summary(wlm_three)
```

#make table

```{r}
table_random <- broom::tidy(wlm_random) %>% 
  mutate(spatial_dist = "random")

table_poisson <- broom::tidy(wlm_poisson) %>% 
  mutate(spatial_dist = "poisson")

table_os <- broom::tidy(wlm_os)%>% 
  mutate(spatial_dist = "matern + 1S")

table_os_om <- broom::tidy(wlm_os_om) %>% 
  mutate(spatial_dist = "matern + 1S + 1M")

table_os_om_two <- broom::tidy(wlm_os_om_two) %>% 
  mutate(spatial_dist = "matern + 1S + 1M + 2")

table_os_two <- broom::tidy(wlm_os_two) %>% 
  mutate(spatial_dist = "matern + 1S + 2")

table_three <- broom::tidy(wlm_three) %>% 
  mutate(spatial_dist = "matern + 3")

model_table <- rbind(table_random, table_poisson, table_os, table_os_om, table_os_om_two, table_os_two, table_three) %>%
  mutate_all(~if(is.numeric(.)) round(., 3) else . )

```

#make predictions and plot

```{r}
#make density data for predictions
true_mussel_density = seq(from=0,to=7,length.out=1000)

#make predictions
p_random = cbind(true_mussel_density, predict(wlm_random, newdata=list(true_mussel_density), interval = "confidence"))

p_poisson = cbind(true_mussel_density,predict(wlm_poisson, newdata=list(true_mussel_density), interval = "confidence"))

p_os = cbind(true_mussel_density, predict(wlm_os, newdata=list(true_mussel_density), interval = "confidence"))

p_os_om = cbind(true_mussel_density,predict(wlm_os_om, newdata=list(true_mussel_density), interval = "confidence"))

p_os_om_two = cbind(true_mussel_density,predict(wlm_os_om_two, newdata=list(true_mussel_density), interval = "confidence"))

p_os_two = cbind(true_mussel_density,predict(wlm_os_two, newdata=list(true_mussel_density), interval = "confidence"))

p_three = cbind(true_mussel_density, predict(wlm_three, newdata=list(true_mussel_density), interval = "confidence"))
```

# Individual plot

```{r}
#set common limits for y scale
y_limits <- c(0, 300)
              
random_plot <- ggplot() +
  geom_point(data = random, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_random, 
            aes(x = true_mussel_density, y = fit), 
            color = "#CC79A7", size = 1) +
  ylim(y_limits) +
  labs(title = "Random",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
  
poisson_plot <- ggplot() +
  geom_point(data = poisson, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_poisson, 
            aes(x = true_mussel_density, y = fit), 
            color = "#D55E00", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Poisson",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

os_plot <- ggplot() +
  geom_point(data = os_matern, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_os, 
            aes(x = true_mussel_density, y = fit), 
            color = "#E69F00", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Matern 1S",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

os_om_plot <- ggplot() +
  geom_point(data = os_om_matern, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_os_om, 
            aes(x = true_mussel_density, y = fit), 
            color = "#56B4E9", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Matern 1S+1M",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

os_om_two_plot <- ggplot() +
  geom_point(data = os_om_two_matern, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_os_om_two, 
            aes(x = true_mussel_density, y = fit), 
            color = "#009E73", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Matern 1S+1M+2",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

os_two_plot <- ggplot() +
  geom_point(data = os_two_matern, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_os_two, 
            aes(x = true_mussel_density, y = fit), 
            color = "#F0E442", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Matern 1S+2",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

three_plot <- ggplot() +
  geom_point(data = three_matern, 
             aes(x = true_mussel_density, y = mussels_per_person_hour), 
             color = "darkgray", alpha = 0.5) + 
  geom_line(data = p_three, 
            aes(x = true_mussel_density, y = fit), 
            color = "#0072B2", size = 1) +
  ylim(y_limits) +
  labs(title = "Clumped Matern 3",
       x = expression(paste("true mussel density (", m^{2}, ")")),
       y = "CPUE") +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
  
```

# Cowplot of all plots
```{r}
# Combine plots into a grid
combined_plot <- plot_grid(
  random_plot, poisson_plot, os_plot, 
  os_om_plot, os_om_two_plot, os_two_plot, 
  three_plot,
  ncol = 3,
  align = "v",
  labels = "AUTO")

ggdraw(combined_plot)

ggsave("figures/H1_combined_plot_detect_0.3.png", plot = combined_plot, width = 12, height = 8)

```


# Combined plot

```{r}
#bind predictions into dataframe
predictions <- data.frame(random = data.frame(p_random)$fit,
                          poisson = data.frame(p_poisson)$fit,
                          os = data.frame(p_os)$fit,
                          os_om = data.frame(p_os_om)$fit,
                          os_om_two = data.frame(p_os_om_two)$fit,
                          os_two = data.frame(p_os_two)$fit,
                          three = data.frame(p_three)$fit,
                          true_mussel_density = true_mussel_density) %>% 
  pivot_longer(!true_mussel_density, names_to = "spatial_distribution", values_to = "CPUE")

#plot predictions 
plot <- ggplot(data = predictions) +
  geom_line(aes(x = true_mussel_density, y = CPUE, color = spatial_distribution), size = 1) +
  labs(x = expression(paste("true mussel density (", m^{2}, ")"))) +
  scale_color_manual(name = "Mussel Spatial Distribution",
                     values = c("random" = "#CC79A7", 
                                "poisson" = "#D55E00",
                                "os" = "#E69F00",
                                "os_om" =  "#56B4E9",
                                "os_om_two" = "#009E73",
                                "os_two" = "#F0E442",
                                "three" = "#0072B2"),
                     labels = c("random" = "Random", 
                                "poisson" = "Poisson",
                                "os" = "Clumped Matern 1S",
                                "os_om" = "Clumped Matern 1S+1M",
                                "os_om_two" = "Clumped Matern 1S+1M+2",
                                "os_two" = "Clumped Matern 1S+2", 
                                "three" = "Clumped Matern 3"),
                     limits = c("random", "poisson", "os", 
                                "os_om", "os_om_two", "os_two", "three")) +
  theme_minimal() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

plot

ggsave("figures/H1_cpue_density_0.3.png", plot = plot, width = 10, height = 8)

```

# Looking at variance


```{r}
#combine all data
combined_data <- rbind(
  random,
  poisson,
  matern
) %>% 
  #get matern labels into spatial dist column
  mutate(spatial_distribution = case_when(spatial_distribution == "Clumped-Matern" ~
                                            matern_clump_placement, 
                                          T ~ spatial_distribution))

#relabel spatial distribution and set factor levels
combined_data$spatial_distribution <- factor(combined_data$spatial_distribution,
                                             levels = c("random", 
                                                        "Clumped-Poisson", 
                                                        "One shore only", 
                                                        "One shore; one middle", 
                                                        "One shore; one middle; 2 additional", 
                                                        "One shore; 2 additional", 
                                                        "Randomly placed"), 
                                             labels = c("Random", 
                                                        "Poisson", 
                                                        "Clumped Matern 1S",
                                                        "Clumped Matern 1S+1M",
                                                        "Clumped Matern 1S+1M+2",
                                                        "Clumped Matern 1S+2",
                                                        "Clumped Matern 3"
                                             )) 
```

Here I am looking at variance. Since high mussel density produces the most variance, I will compare variance between spatial distributions when mussel density is 7 mussels per meter.

- Variance

- Standard Deviation

- Interquartile Range

- Range

- Coefficient of Variation

- Skewness

- Kurtosis 

- Outliers

## Data Viz

```{r}
library(gganimate)
library(transformr)

combined_data <- combined_data %>% filter(true_mussel_density == 7)

# Violin plot to show distribution and variation
vio_plot <- ggplot(combined_data, aes(x = spatial_distribution, y = mussels_per_person_hour, 
                          fill = spatial_distribution)) +
  geom_violin(scale = "count", 
              trim = T) + 
   scale_fill_manual(name = "Mussel Spatial Distribution",
                     values = c("Random" = "#CC79A7", 
                                "Poisson" = "#D55E00",
                                "Clumped Matern 1S"= "#E69F00",
                                "Clumped Matern 1S+1M" =  "#56B4E9",
                                "Clumped Matern 1S+1M+2" = "#009E73",
                                "Clumped Matern 1S+2" = "#F0E442",
                                "Clumped Matern 3" = "#0072B2")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_minimal() +
  labs(title = paste("Distribution of CPUE when mussel density is 7 mussels per meter"), 
       x = "Spatial Distribution", 
       y = "CPUE") +
  theme(legend.position = "none",
        axis.title.y = element_text(angle = 0, vjust = 0.5)
        ) + 
  ylim(0, NA)

#print it
vio_plot

#animate the plot
# animated_plot <- vio_plot + 
#   transition_states(
#     true_mussel_density, 
#     transition_length = 2, 
#     state_length = 1, 
#     wrap = TRUE
#   ) +
#   enter_fade() + 
#   exit_fade() +
#   ease_aes('linear')

# #render it
# rendered_anim <- animate(animated_plot, nframes = 100, fps = 10, width = 600, height = 400, renderer = gifski_renderer())
# 
# anim_save("figures/violin_plot_cpue_by_density.gif", rendered_anim)

ggsave("figures/H1_distribution_cpue_density_7.png", plot = vio_plot, width = 10, height = 8)
```

## SD and Range

```{r}
combined_variance <- combined_data %>% 
  group_by(spatial_distribution, true_mussel_density) %>% 
  summarise(min = min(mussels_per_person_hour),
            max = max(mussels_per_person_hour),
            mean = mean(mussels_per_person_hour),
            variance = var(mussels_per_person_hour), 
            standard_dev = sd(mussels_per_person_hour)) %>% 
   mutate_if(is.numeric, round, digits = 2)
```


```{r}
ggplot(data = combined_variance) +
  geom_line(aes(x = true_mussel_density, y = mean, color = spatial_distribution)) +
  labs(title = "Mean CPUE by spatial distribution as density increases")
```


```{r}
ggplot(data = combined_variance) +
  geom_line(aes(x = true_mussel_density, y = standard_dev, color = spatial_distribution)) +
  labs(title = "Standard deviation in CPUE by spatial distribution as density increases")

```



# Density from Timed Search

```{r}
ggplot(data = random) +
  geom_point(aes(x = true_mussel_density, y = (total_mussels_detected/(total_quadrats_sampled/4))))

ggplot(data = os_matern) +
  geom_point(aes(x = true_mussel_density, y = (total_mussels_detected/(total_quadrats_sampled/4))))
```

