---
title: "H3 Curve Fitting"
author: "Anonymous"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
toc-title: "Table of contents"
sansfont: Calibri
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(janitor)
library(kableExtra)
library(cowplot)
```

# Defining Modeling Functions

## functions to fit different types of curves

```{r warning=FALSE, message=FALSE}
#exponential decay
fit_exponential <- function(df, 
                            residual_col, 
                            intensity_col) {
  
  #extract relevant cols
  residual <- df[[residual_col]]
  intensity <- df[[intensity_col]]
  
  #define model formula
  model_formula <- residual ~ a * exp(-b * (intensity))
  
  #fit the model
  model<- tryCatch({
    nls(model_formula,
        start = list(a = max(residual), b = 0.01))
  }, error = function(e) {
    message("Model fitting failed: ", e$message)
  })
  
  return(model)
}

#exponential decay plus constant
fit_decay <- function(df, residual_col, intensity_col) {
  
  residual <- df[[residual_col]]
  intensity <- df[[intensity_col]]
  
  #define the model formula
  model_formula <- residual ~ a * exp(-b * intensity) + c
  
  #fit the model
  model <- tryCatch(
    nls(model_formula, 
        data = df, 
        start = list(a = 1, b = 0.01, c = 0)),
    error = function(e) {
      message("Model fitting failed: ", e$message)
    })
  
  return(model)
}

#power law decay
fit_power_law <- function(df, residual_col, intensity_col) {
  
  residual <- df[[residual_col]]
  intensity <- df[[intensity_col]]
  
  #define power law decay formula
  #model_formula <- residual ~ a * intensity^(-b) + c
  model_formula <- residual ~ a * intensity^(-b)
  
  #list start vals
  start_vals <- list(
    a = max(df$residual, na.rm = TRUE),
    b = 1
    #c = 0
  )
  
  #fit the model
  model <- tryCatch(
    nls(model_formula,
        data = df, 
        start = start_vals),
    error = function(e) {
      message("Model fitting failed: ", e$message)
    })
  
  return(model)
}

#log transformed linear model
fit_log <- function(df, residual_col, intensity_col) {
  
  residual <- df[[residual_col]]
  intensity <- df[[intensity_col]]
  
  #fit the model
  model <- tryCatch(
    lm(residual ~ log(intensity)), #this version for timed search
    error = function(e) {
      message("Model fitting failed: ", e$message)
    })
  
  return(model)
}

#linear model
fit_line <- function(df, residual_col, intensity_col) {
  
  residual <- df[[residual_col]]
  intensity <- df[[intensity_col]]
  
  #fit the model
  model <- tryCatch(
    lm(residual ~ intensity), 
    error = function(e) {
      message("Model fitting failed: ", e$message)
    })
  
  return(model)
}
```

## functions to make predictions and summarise models

```{r warning=FALSE, message=FALSE}
#function to predict values using the fitted model
exp_predict <- function(df, new_df, model_type = c("exp", "pwr", "decay", "log", "line", "ploy"), residual_col, intensity_col) {
  
  #fit the correct model
  if(model_type == "exp") {
    model <- fit_exponential(df, residual_col, intensity_col)
  } 
  if(model_type == "pwr") {
    model <- fit_power_law(df, residual_col, intensity_col)
  } 
  if(model_type == "decay") {
    model <- fit_decay(df, residual_col, intensity_col)
  }
  if(model_type == "log") {
    model <- fit_log(df, residual_col, intensity_col)
  }
  if(model_type == "line") {
    model <- fit_line(df, residual_col, intensity_col)
  }
  if(model_type == "poly") {
    model <- fit_poly(df, residual_col, intensity_col)
  }
  
  #if model fits, generate predictions df
  if (!is.null(model)) {
    pred_df <- data.frame(
      intensity = new_df$intensity,
      predicted = predict(model, newdata = new_df),
      rse = summary(model)$sigma
    )
    return(pred_df)
  } else {
    failed_df <- new_df %>%
      mutate(group = NA) %>% 
      mutate(predicted = NA) %>% 
      mutate(rese = NA)
    
    return(failed_df)
  }
}

#function to print estimated coefficients
exp_coeff <- function(df, model_type = c("exp", "pwr", "decay", "log", "line", "poly"), residual_col, intensity_col) {
  
  if(model_type == "exp") {
    model <- fit_exponential(df, residual_col, intensity_col)
  } 
  if(model_type == "pwr") {
    model <- fit_power_law(df, residual_col, intensity_col)
  } 
  if(model_type == "decay") {
    model <- fit_decay(df, residual_col, intensity_col)
  }
  if(model_type == "log") {
    model <- fit_log(df, residual_col, intensity_col)
  }
  if(model_type == "line") {
    model <- fit_line(df, residual_col, intensity_col)
  }
  if(model_type == "poly") {
    model <- fit_poly(df, residual_col, intensity_col)
  }
  
  #if the model fits, return tidy model summary and rse
  if (!is.null(model)) {
    tidy_summary <- broom::tidy(model) %>% 
      mutate(rse = summary(model)$sigma)
    return(tidy_summary)
  } else {
    return(data.frame(term = c(NA), estimate = (NA), std.error = (NA), statistic = c(NA), p.value = c(NA), rse = c(NA)))
  }
}

#function to print model goodness of fit info
model_info <- function(df, model_type = c("exp", "pwr", "decay", "log", "line", "poly"), residual_col, intensity_col) {
  
  if(model_type == "exp") {
    model <- fit_exponential(df, residual_col, intensity_col)
  } 
  if(model_type == "pwr") {
    model <- fit_power_law(df, residual_col, intensity_col)
  } 
  if(model_type == "decay") {
    model <- fit_decay(df, residual_col, intensity_col)
  }
  if(model_type == "log") {
    model <- fit_log(df, residual_col, intensity_col)
  }
  if(model_type == "line") {
    model <- fit_line(df, residual_col, intensity_col)
  }
  if(model_type == "poly") {
    model <- fit_poly(df, residual_col, intensity_col)
  }
  
  if (!is.null(model)) {
    model_info <- data.frame(
      rse = summary(model)$sigma
    )
    return(model_info)
    
  } else {
    model_info <- data.frame(
      rse = c(NA)
    )
    return(model_info)
  }
}

```

# Reading in Data

```{r warning=FALSE, message=FALSE}
#matern distribution data
h3_matern_acs <- read.csv("data/H3_matern_acs.csv")
h3_matern_srs <- read.csv("data/H3_matern_srs.csv")
h3_matern_timed <- read.csv("data/H3_matern_timed.csv")
h3_matern_transect <- read.csv("data/H3_matern_transect.csv")

#poisson distribution data
h3_poisson_acs <- read.csv("data/H3_poisson_acs.csv") %>% 
  filter(Number.of.clumps == 5, Poisson.mean == 3)
h3_poisson_srs <- read.csv("data/H3_poisson_srs.csv") %>% 
  filter(Number.of.clumps == 5, Poisson.mean == 3)
h3_poisson_timed <- read.csv("data/H3_poisson_timed.csv") %>% 
  filter(Number.of.clumps == 5, Poisson.mean == 3)
h3_poisson_transect <- read.csv("data/H3_poisson_transect.csv") %>% 
  filter(Number.of.clumps == 5, Poisson.mean == 3)

#random distribution data
h3_random_acs <- read.csv("data/H3_random_acs.csv")
h3_random_srs <- read.csv("data/H3_random_srs.csv")
h3_random_timed <- read.csv("data/H3_random_timed.csv")
h3_random_transect <- read.csv("data/H3_random_transect.csv")

#create new data for predictions
new_dat <- data.frame(intensity = c(seq(0, 500, length.out = 100)))
```


# SRS

clean up the column names and compute absolute value of the residual

```{r warning=FALSE, message=FALSE}
#clean_names
srs_clean <- rbind(h3_matern_srs, h3_poisson_srs, h3_random_srs) %>% 
  janitor::clean_names() %>% 
  mutate(spatial_distribution = case_when(spatial_distribution == "Clumped-Matern" ~ paste(matern_clump_placement), T ~ spatial_distribution)) %>% 
  filter(detectability_of_rare_mussels == 1)

#calculate residual
srs_residual <- srs_clean %>% 
  #mutate(residual = abs(true_mussel_density - estimated_density)) %>%
  mutate(residual = (abs(estimated_density-true_mussel_density)/true_mussel_density) * 100) %>%
  mutate(m_sampled = total_quadrats_sampled/4) %>% 
  mutate(density_cat = case_when(true_mussel_density %in% c(0.1, 0.01) ~ "rare", T ~ "common"))
```

## Quick check of relationship

```{r}
srs_mean <- srs_residual %>% 
  group_by(m_sampled, spatial_distribution, density_cat) %>% 
  summarise(mean_pe = mean(residual))

ggplot(data = srs_mean) +
  geom_point(aes(y = mean_pe, x = m_sampled, 
                 color = spatial_distribution, shape = density_cat)) +
  geom_line(aes(x = m_sampled, y = mean_pe, colour = spatial_distribution, linetype = density_cat))

```


## Exponential model results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
srs_exp_model_summary <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
srs_exp_model_coeff <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
srs_exp_model_preds <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "exp", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(srs_exp_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = as.factor(density_cat),
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "SRS: Exponential Decay of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()

```

## Power Law Decay Results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
srs_pwr_model_summary <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
srs_pwr_model_coeff <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
srs_pwr_model_preds <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "pwr", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(srs_pwr_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = as.factor(density_cat),
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "SRS: Power Law Decay of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()

```
NOTES: Power law decay fit the mattern and random distributions at some densities. In all cases when the power law fit, it had a marginally lower RSE than the basic exponential decay model.

## Exponential Decay + constant results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
srs_dec_model_summary <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
srs_dec_model_coeff <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
srs_dec_model_preds <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "decay", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(srs_dec_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = as.factor(density_cat),
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "SRS: Decay  + Constant Model of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Log transformed linear model

```{r warning=FALSE, message=FALSE}

#evaluate model fit
srs_log_model_summary <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
srs_log_model_coeff <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
srs_log_model_preds <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "log", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting

# Plot the data and fitted curves
ggplot(srs_log_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = as.factor(density_cat),
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "SRS: Log Transformed LM of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Linear

```{r warning=FALSE, message=FALSE}

#evaluate model fit
srs_line_model_summary <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
srs_line_model_coeff <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
srs_line_model_preds <- srs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "line", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting

# Plot the data and fitted curves
ggplot(srs_residual, aes(x = m_sampled, y = residual)) +
  #geom_point(aes(color = spatial_distribution))  +
  geom_line(data = srs_line_model_preds, aes(x = intensity, y = predicted,
                                             linetype = as.factor(density_cat),
                                             color = spatial_distribution),
            linewidth = 1) +
  labs(title = "SRS: LM of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```


## Table of RSEs

```{r warning=FALSE, message=FALSE}
#bind RSE into table
srs_rse_table <- data.frame(
  spatial_distribution = c(srs_log_model_summary$spatial_distribution),
  density = c(srs_log_model_summary$density_cat),
  exp_rse = c(srs_exp_model_summary$rse),
  pwr_rse = c(srs_pwr_model_summary$rse),
  decay_rse = c(srs_dec_model_summary$rse),
  log_rse = c(srs_log_model_summary$rse),
  line_rse = c(srs_line_model_summary$rse)) 

#add colum which has min rse model name
y <- colnames(srs_rse_table)[apply(srs_rse_table[,c(3:7)],1,which.min) + 2]
srs_rse_table$min_col <- y

knitr::kable(srs_rse_table) %>% 
  kable_styling()
```

## Best Models Plotted Together

```{r}
#bind RSE into table
srs_best_fit_preds <- rbind(
  srs_exp_model_preds,
  srs_dec_model_preds,
  srs_pwr_model_preds,
  srs_log_model_preds
) %>% 
  group_by(spatial_distribution, density_cat, intensity) %>% 
  slice_min(order_by = rse, n = 1, with_ties = T) %>% 
  mutate(search_method = "SRS")

# Plot the data and fitted curves
srs_bestfit <- 
  ggplot() +
  geom_line(data = srs_best_fit_preds, aes(x = intensity, y = predicted,
                                           linetype = density_cat,
                                           color = spatial_distribution),
            linewidth = 1) +
  labs(title = "Simple Random Sample",
       x = expression(paste("Area Sampled (", m^{2}, ")")),
       y = "Percent Error") +
  scale_color_manual(values = c("random" = "#CC79A7", 
                                "Clumped-Poisson" = "#D55E00",
                                "One shore only" = "#E69F00",
                                "One shore; one middle" = "#56B4E9",
                                "One shore; one middle; 2 additional" = "#009E73",
                                "One shore; 2 additional" = "#F0E442",
                                "Randomly placed" = "#0072B2"),
                     labels = c("random" = "Random", 
                                "Clumped-Poisson" = "Poisson",
                                "One shore only" = "Clumped Matern 1S",
                                "One shore; one middle" = "Clumped Matern 1S+1M",
                                "One shore; one middle; 2 additional" = "Clumped Matern 1S+1M+2",
                                "One shore; 2 additional" = "Clumped Matern 1S+2", 
                                "Randomly placed" = "Clumped Matern 3"),
                     limits = c("random", 
                                "Clumped-Poisson", 
                                "One shore only", 
                                "One shore; one middle", 
                                "One shore; one middle; 2 additional", 
                                "One shore; 2 additional", 
                                "Randomly placed")) +
  scale_y_continuous(limits = c(0, 200)) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

```

# Transects

```{r warning=FALSE, message=FALSE}
#clean_names
transect_clean <- rbind(h3_matern_transect, h3_poisson_transect, h3_random_transect) %>% 
  janitor::clean_names() %>% 
  mutate(spatial_distribution = case_when(spatial_distribution == "Clumped-Matern" ~ paste(matern_clump_placement), T ~ spatial_distribution)) %>% 
  filter(detectability_of_rare_mussels == 1)

#calculate residual
transect_residual <- transect_clean %>% 
  mutate(residual = (abs(estimated_density-true_mussel_density)/true_mussel_density) * 100) %>%
  mutate(m_sampled = total_quadrats_sampled/4) %>% 
  mutate(density_cat = case_when(true_mussel_density %in% c(0.1, 0.01) ~ "rare", T ~ "common"))
```

```{r}
transect_mean <- transect_residual %>% 
  group_by(m_sampled, spatial_distribution, density_cat) %>% 
  summarise(mean_pe = mean(residual))

ggplot(data = transect_mean) +
  geom_point(aes(y = mean_pe, x = m_sampled, 
                 color = spatial_distribution, shape = density_cat)) +
  geom_line(aes(x = m_sampled, y = mean_pe, colour = spatial_distribution, linetype = density_cat))

```


## Exponential model results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
transect_exp_model_summary <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
transect_exp_model_coeff <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
transect_exp_model_preds <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "exp", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(transect_exp_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "transect: Exponential Decay of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()

```

## Power Law Decay Results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
transect_pwr_model_summary <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
transect_pwr_model_coeff <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
transect_pwr_model_preds <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "pwr", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(transect_pwr_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "transect: Power Law Decay of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()

```

## Exponential Decay + constant results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
transect_dec_model_summary <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))


#get coefficient estimates
transect_dec_model_coeff <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
transect_dec_model_preds <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "decay", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(transect_dec_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "transect: Decay  + Constant Model of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Log transformed linear model

```{r warning=FALSE, message=FALSE}
#evaluate model fit
transect_log_model_summary <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
transect_log_model_coeff <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
transect_log_model_preds <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "log", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting


# Plot the data and fitted curves
ggplot(transect_log_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "transect: Log Transformed LM of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```


## Linear

```{r warning=FALSE, message=FALSE}

#evaluate model fit
transect_line_model_summary <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
transect_line_model_coeff <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
transect_line_model_preds <- transect_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "line", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting

# Plot the data and fitted curves
ggplot(transect_line_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "transect: LM of Residual Error as Sampling Intensity Increases",
       x = "Square m Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Table of RSEs

```{r warning=FALSE, message=FALSE}
#bind RSE into table
transect_rse_table <- data.frame(
  spatial_distribution = c(transect_log_model_summary$spatial_distribution),
  density = c(transect_log_model_summary$density_cat),
  exp_rse = c(transect_exp_model_summary$rse),
  pwr_rse = c(transect_pwr_model_summary$rse),
  decay_rse = c(transect_dec_model_summary$rse),
  log_rse = c(transect_log_model_summary$rse),
  line_rse = c(transect_line_model_summary$rse))

#add column which has min rse model name
y <- colnames(transect_rse_table)[apply(transect_rse_table[,c(3:7)],1,which.min) + 2]
transect_rse_table$min_col <- y

knitr::kable(transect_rse_table) %>% 
  kable_styling()
```

## Best Models Plotted Together

```{r}
#bind RSE into table
trs_best_fit_preds <- rbind(
  transect_exp_model_preds,
  transect_dec_model_preds,
  transect_pwr_model_preds,
  transect_log_model_preds,
) %>% 
  group_by(spatial_distribution, density_cat, intensity) %>% 
  slice_min(order_by = rse, n = 1, with_ties = T) %>% 
  mutate(search_method = "TRS")

# Plot the data and fitted curves
trs_bestfit <- 
  ggplot() +
  geom_line(data = trs_best_fit_preds, aes(x = intensity, y = predicted,
                                           linetype = density_cat,
                                           color = spatial_distribution),
            linewidth = 1) +
  labs(title = "Transect Sample",
       x = expression(paste("Area Sampled (", m^{2}, ")")),
       y = "Percent Error") +
  scale_color_manual(values = c("random" = "#CC79A7", 
                                "Clumped-Poisson" = "#D55E00",
                                "One shore only" = "#E69F00",
                                "One shore; one middle" = "#56B4E9",
                                "One shore; one middle; 2 additional" = "#009E73",
                                "One shore; 2 additional" = "#F0E442",
                                "Randomly placed" = "#0072B2"),
                     labels = c("random" = "Random", 
                                "Clumped-Poisson" = "Poisson",
                                "One shore only" = "Clumped Matern 1S",
                                "One shore; one middle" = "Clumped Matern 1S+1M",
                                "One shore; one middle; 2 additional" = "Clumped Matern 1S+1M+2",
                                "One shore; 2 additional" = "Clumped Matern 1S+2", 
                                "Randomly placed" = "Clumped Matern 3"),
                     limits = c("random", 
                                "Clumped-Poisson", 
                                "One shore only", 
                                "One shore; one middle", 
                                "One shore; one middle; 2 additional", 
                                "One shore; 2 additional", 
                                "Randomly placed")) +
  scale_y_continuous(limits = c(0, 200)) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

```

# ACS

```{r warning=FALSE, message=FALSE}
#clean_names
acs_clean <- rbind(h3_matern_acs, h3_poisson_acs, h3_random_acs) %>% 
  janitor::clean_names() %>% 
  mutate(spatial_distribution = case_when(spatial_distribution == "Clumped-Matern" ~ paste(matern_clump_placement), T ~ spatial_distribution)) %>% 
  filter(detectability_of_rare_mussels == 1)

#calculate residual
acs_residual <- acs_clean %>% 
  mutate(residual = (abs(hh_metric-true_mussel_density)/true_mussel_density) * 100) %>%
  mutate(m_sampled = total_quadrats_sampled/4) %>% 
  mutate(density_cat = case_when(true_mussel_density %in% c(0.1, 0.01) ~ "rare", T ~ "common"))
```

```{r}
acs_mean <- acs_residual %>% 
  group_by(m_sampled, spatial_distribution, density_cat) %>% 
  summarise(mean_pe = mean(residual))

#note try starting quadrats and total area sampled
ggplot(data = acs_mean) +
  geom_point(aes(y = mean_pe, x = m_sampled, 
                 color = spatial_distribution, shape = density_cat)) +
  geom_line(aes(x = m_sampled, y = mean_pe, colour = spatial_distribution, linetype = density_cat)) +
  scale_y_continuous(limits = c(0, 200)) 

```

## Exponential model results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
acs_exp_model_summary <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
acs_exp_model_coeff <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "exp", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
acs_exp_model_preds <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "exp", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(acs_exp_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "acs: Exponential Decay of Residual Error as Sampling Intensity Increases",
       x = "Total Meters sampled",
       y = "Residual Error") +
  theme_minimal()

```

## Power Law Decay Results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
acs_pwr_model_summary <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
acs_pwr_model_coeff <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "pwr", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
acs_pwr_model_preds <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "pwr", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(acs_pwr_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "acs: Power Law Decay of Residual Error as Sampling Intensity Increases",
       x = "Total Meters sampled",
       y = "Residual Error") +
  theme_minimal()

```

## Exponential Decay + constant results

```{r warning=FALSE, message=FALSE}
#evaluate model fit
acs_dec_model_summary <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
acs_dec_model_coeff <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "decay", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
acs_dec_model_preds <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "decay", residual = "residual", intensity = "m_sampled"))

# Plot the data and fitted curves
ggplot(acs_dec_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "acs: Decay  + Constant Model of Residual Error as Sampling Intensity Increases",
       x = "Total Meters Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Log transformed linear model

```{r warning=FALSE, message=FALSE}
#evaluate model fit
acs_log_model_summary <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
acs_log_model_coeff <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "log", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
acs_log_model_preds <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "log", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting


# Plot the data and fitted curves
ggplot(acs_log_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "acs: Log Transformed LM of Residual Error as Sampling Intensity Increases",
       x = "Total Meters Sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Linear

```{r warning=FALSE, message=FALSE}

#evaluate model fit
acs_line_model_summary <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ model_info(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get coefficient estimates
acs_line_model_coeff <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify(~ exp_coeff(.x, model_type = "line", residual_col = "residual", intensity_col = "m_sampled"))

#get 100 predictions per group
acs_line_model_preds <- acs_residual %>%
  group_by(spatial_distribution, density_cat) %>%
  group_modify( ~ exp_predict(.x, new_df = new_dat, model_type = "line", residual = "residual", intensity = "m_sampled")) %>% 
  filter(!is.infinite(predicted)) #remove infinite vals for plotting

# Plot the data and fitted curves
ggplot(acs_line_model_preds, aes(x = intensity, y = predicted)) +
  geom_line(aes(linetype = density_cat,
                color = spatial_distribution),
            linewidth = 1) +
  labs(title = "acs: LM of Residual Error as Sampling Intensity Increases",
       x = "m_sampled",
       y = "Residual Error") +
  theme_minimal()
```

## Table of RSEs

```{r warning=FALSE, message=FALSE}
#bind RSE into table
acs_rse_table <- data.frame(
  spatial_distribution = c(acs_log_model_summary$spatial_distribution),
  density = c(acs_log_model_summary$density_cat),
  exp_rse = c(acs_exp_model_summary$rse),
  pwr_rse = c(acs_pwr_model_summary$rse),
  decay_rse = c(acs_dec_model_summary$rse),
  log_rse = c(acs_log_model_summary$rse),
  line_rse = c(acs_line_model_summary$rse))

#add column which has min rse model name
y <- colnames(acs_rse_table)[apply(acs_rse_table[,c(3:7)],1,which.min) + 2]
acs_rse_table$min_col <- y

knitr::kable(acs_rse_table) %>% 
  kable_styling()
```

## Best Models Plotted Together

```{r}
#bind RSE into table
acs_best_fit_preds <- rbind(
  acs_exp_model_preds,
  acs_dec_model_preds,
  acs_pwr_model_preds,
  acs_log_model_preds
) %>% 
  group_by(spatial_distribution, density_cat, intensity) %>% 
  slice_min(order_by = rse, n = 1, with_ties = T) %>% 
  mutate(search_method = "ACS")

# Plot the data and fitted curves
acs_bestfit <- 
  ggplot() +
  geom_line(data = acs_best_fit_preds, aes(x = intensity, y = predicted,
                                           color = spatial_distribution,
                                           linetype = density_cat),
            linewidth = 1) +
  scale_y_continuous(limits = c(0, 200)) +
  labs(title = "Adaptive Cluster Sample",
       x = expression(paste("Area Sampled (", m^{2}, ")")),
       y = "Percent Error") +
  scale_color_manual(name = "Spatial Distribution",
values = c("random" = "#CC79A7", 
                                "Clumped-Poisson" = "#D55E00",
                                "One shore only" = "#E69F00",
                                "One shore; one middle" = "#56B4E9",
                                "One shore; one middle; 2 additional" = "#009E73",
                                "One shore; 2 additional" = "#F0E442",
                                "Randomly placed" = "#0072B2"),
                     labels = c("random" = "Random", 
                                "Clumped-Poisson" = "Poisson",
                                "One shore only" = "Clumped Matern 1S",
                                "One shore; one middle" = "Clumped Matern 1S+1M",
                                "One shore; one middle; 2 additional" = 
                                  "Clumped Matern 1S+1M+2",
                                "One shore; 2 additional" = "Clumped Matern 1S+2", 
                                "Randomly placed" = "Clumped Matern 3"),
                     limits = c("random", 
                                "Clumped-Poisson", 
                                "One shore only", 
                                "One shore; one middle", 
                                "One shore; one middle; 2 additional", 
                                "One shore; 2 additional", 
                                "Randomly placed")) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))

```

# Cowplot of Best Plots

```{r}
color_legend <- ggplot() +
  geom_line(data = acs_best_fit_preds, aes(x = intensity, y = predicted,
                                             color = spatial_distribution),
            linewidth = 1) +
  scale_color_manual(name = "Mussel Spatial Distribution",
                     values = c("random" = "#CC79A7", 
                                "Clumped-Poisson" = "#D55E00",
                                "One shore only" = "#E69F00",
                                "One shore; one middle" = "#56B4E9",
                                "One shore; one middle; 2 additional" = "#009E73",
                                "One shore; 2 additional" = "#F0E442",
                                "Randomly placed" = "#0072B2"),
                     labels = c("random" = "Random", 
                                "Clumped-Poisson" = "Poisson",
                                "One shore only" = "Clumped Matern 1S",
                                "One shore; one middle" = "Clumped Matern 1S+1M",
                                "One shore; one middle; 2 additional" = "Clumped Matern 1S+1M+2",
                                "One shore; 2 additional" = "Clumped Matern 1S+2", 
                                "Randomly placed" = "Clumped Matern 3"),
                     limits = c("random", 
                                "Clumped-Poisson", 
                                "One shore only", 
                                "One shore; one middle", 
                                "One shore; one middle; 2 additional", 
                                "One shore; 2 additional", 
                                "Randomly placed")) +
  theme_minimal() +
  theme(legend.position = "right")


line_legend <- ggplot() +
  geom_line(data = acs_best_fit_preds, aes(x = intensity, y = predicted,
                                             linetype = density_cat),
            linewidth = 1) +
  labs(title = "Timed Search",
       x = "Number of Person-Hours Searched",
       y = "Proportion Found") +
  scale_linetype_manual(name = expression(
    paste("True Mussel Density")),
    values = c("rare" = "dashed", "common" = "solid")) +
  theme_minimal() +
  theme(legend.position = "right")
```


```{r}

get_legend_35 <- function(plot) {
  # return all legend candidates
  legends <- get_plot_component(plot, "guide-box", return_all = TRUE)
  # find non-zero legends
  nonzero <- vapply(legends, \(x) !inherits(x, "zeroGrob"), TRUE)
  idx <- which(nonzero)
  # return first non-zero legend if exists, and otherwise first element (which will be a zeroGrob) 
  if (length(idx) > 0) {
    return(legends[[idx[1]]])
  } else {
    return(legends[[1]])
  }
}

# Extract the legend from one of the plots
c_legend <- get_legend_35(color_legend + 
                              guides(color = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom"))

l_legend <- get_legend_35(line_legend + 
                              guides(color = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom"))

# Combine the plots into a 2x2 grid
combined_plot <- plot_grid(srs_bestfit, trs_bestfit, acs_bestfit, 
                           ncol = 3,  # Number of columns
                           align = "v",  # Align vertically
                           labels = "AUTO")  # Automatically label the plots

# Combine legends into a single plot
combined_legend <- plot_grid(c_legend, l_legend, nrow = 2, align = 'h', rel_heights = c(1, 1))


# Combine plots with the legend
final_plot <- plot_grid(combined_plot, combined_legend, ncol = 1, rel_heights = c(4, 1))

# Print the combined plot
ggdraw(final_plot)

ggsave("figures/H3_combined_plot_PE.png", plot = final_plot, width = 12, height = 4)
```

```{r}
#saving best predictions
pe_predictions <- rbind(srs_best_fit_preds, acs_best_fit_preds, trs_best_fit_preds)

write.csv(pe_predictions, "data/pe_predictions.csv")
```


