---
title: "Mussel Sampling Model Variance Exploration"
author: "Anonymous"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(janitor)
```

# Purpose 

The purpose of this exploration is to determine how many model runs are needed per each scenario. To do this, we will examine how many runs are required for variation in estimated density/CPUE to stabilize.

Note: The data used is from the behavior space "variance_convergence" which runs 9600 times. spatial distribution was set to poisson cluster, 10 clusters, poisson mean of 2 for all model runs. Quadrat size was set to 0.5.

```{r}
#read in data (for now copied into R project)
raw_data <- read.csv("data/variance_convergence_6-17-2024.csv", sep = ",")

#clean names
clean_data <- clean_names(raw_data) %>% 
  rename(mussel_density = true_mussel_density) #rename to shorter name
```

```{r}
#write function for calculating cumulative variance 
calc_cum_var <- function (x) {
  n <- length(x)
  cum_var <- numeric(n)
  for (i in 2:n){
    cum_var[i] <- var(x[1:i])
  }
  return(cum_var)
  }
```

# Simple Random Sample (SRS)

In this section we will look at variance in estimated density when the sampling method is simple-random sample. 

```{r}
#filter for only simple random sample
all_srs <- filter(clean_data, sampling_method == "simple-random-sample")
```

For each SRS scenario there are different numbers of quadrats in the SRS and different mussel densities.

```{r}
#group data by mussel density and number of quadrats and calc cumulative variance
var <- all_srs %>% 
  group_by(mussel_density, srs_num_quadrats) %>% 
  mutate(run_id = row_number()) %>% #arbitrary row number becomes x axis
  mutate(variance_est_density = calc_cum_var(estimated_density))

#plot row number on x axis, cumulative variance on y axis
ggplot(data = var, aes(x=run_id, y = variance_est_density)) +
  geom_line() +
  facet_grid(mussel_density ~ srs_num_quadrats, 
             beller=label_bothla, scales = "free_y") +
  labs(title = "Cumulative Variance in Estimated Density (SRS)",
       y = "variance in\nestimated density",
       x = "number of model runs") +
  theme_bw() +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
  
```


# Transects 

In this section we will look at variance in estimated density when the sampling method is Transect Sampling. 

For each transect scenario there are different numbers of transects, different numbers of quadrats sampled in each transect and different mussel densities.


```{r}
#filter for only transect sampling
all_tran <- filter(clean_data, sampling_method == "transect")

#group by mussel density, number of transects, and quadrats per transect
#then calculate the cumulative variance
var_tran <- all_tran %>% 
  group_by(mussel_density, 
           number_transects, 
           number_quadrats_per_transect) %>% 
  mutate(run_id = row_number()) %>% 
  mutate(number_quadrats_per_transect = as.character(number_quadrats_per_transect)) %>% 
  mutate(variance_est_density = calc_cum_var(estimated_density))

#plot the variance across model runs
ggplot(data = var_tran, aes(x=run_id, y = variance_est_density, color = number_quadrats_per_transect)) +
  geom_line() +
  facet_grid(mussel_density ~ number_transects, 
             labeller=label_both, scales = "free_y") +
  labs(title = "Transect Cumulative Variance in Estimated Density")
```


# Adaptive Cluster Sample (ACS)

In this section we will look at variance in estimated density (HH metric) when the sampling method is Adaptive Cluster Sampling. 

For each ACS scenario there are different numbers of initial quadrats in the ACS and different mussel densities.

```{r}
#filter for only ACS
all_acs <- filter(clean_data, sampling_method == "adaptive-cluster")

#group by mussel density and number of initial quadrats
var_acs <- all_acs %>% 
  group_by(mussel_density, acs_num_quadrats) %>% 
  mutate(run_id = row_number()) %>% 
  mutate(variance_hh = calc_cum_var(hh_metric))

#plot cumulative variance in the HH metric across model runs
ggplot(data = var_acs, aes(x=run_id, y = variance_hh)) +
  geom_line() +
  facet_grid(mussel_density ~ acs_num_quadrats, 
             labeller=label_both, scales = "free_y") +
  labs(title = "ACS Cumulative Variance in HH Metric")
```

# Random Walks (evaluated using CPUE)

In this section we will look at variance in catch per unit effort (CPUE - number of mussels detected in one person-hour) when the sampling method is random walking or timed search. 

For each random walk scenario there are different mussel densities.

```{r}
#filter for only random walk sampling
all_walk <- filter(clean_data, sampling_method == "timed-search")

#group by mussel density and calculate cumulative variance
var_walk <- all_walk %>% 
  mutate(mussel_density = as.character(mussel_density)) %>% 
  group_by(mussel_density) %>% 
  mutate(run_id = row_number()) %>% 
  mutate(variance_cpue = calc_cum_var(mussels_per_person_hour))

#plot variance in CPUE
ggplot(data = var_walk, aes(x=run_id, y = variance_cpue, color = mussel_density)) +
  geom_line() +
  labs(title = "Random walk Cumulative Variance in CPUE")
```

